---
title: "R versions"
output: html_document
editor_options:
  chunk_output_type: console
---

**Subject**: Important: R Installation Instructions for 2024_RNA-Seq-Analysis

**Please read this entire message carefully.**

**For those of you who are on a Mac, please install the x86_64 version of R, even if you are using an M1 Mac.** This is important for compatibility with the packages we will be using, especially Bioconductor (e.g. `igraph`).

Here’s how to do it:

1. **Download R**:

- Visit the [CRAN R Project](https://cran.r-project.org/).

- Click on "Download R for (Mac) OS X".

- Select the **R-4.2.3.pkg** (or the latest x86_64 version) for Intel Macs.

2. **Installing R**:

- Make sure you have the latest version of RStudio installed on your computer. You can download it [here](https://www.rstudio.com/products/rstudio/download/).

- Make sure you quit RStudio before installing R.
- Open the downloaded `.pkg` file and follow the installation instructions.

3. **Using Rosetta for those of you using an M1 Mac**: 

- Rosetta 2 enables a Mac with Apple silicon chip M1 to use apps built for a Mac with an Intel processor.

- Rosetta 2 is available only for Mac computers with Apple silicon chip M1.

- Make sure that your Mac is connected to the internet.

- If you are using an M1 Mac, you may need to install Rosetta 2 to run the x86_64 version of R. Here’s how to do it:

- When you first open RStudio, you may see a message asking if you want to install Rosetta 2. Click "Install" to proceed.

- If you have the option, ensure that your terminal or RStudio is running under   Rosetta 2. To do this:

- Right-click on the RStudio icon in the Applications folder. Select **Get Info**.

- Check **Open using Rosetta**

After installation, please check that you are using the x86_64 version of R by running the following command in your R console:

```{r}

R.version

```

It should return something like this:

platform       x86_64-apple-darwin20       
arch           x86_64                      
os             darwin20                    
system         x86_64, darwin20            
status                                     
major          4                           
minor          4.2                         
year           2024                        
month          10                          
day            31                          
svn rev        87279                       
language       R                           
version.string R version 4.4.2 (2024-10-31)
nickname       Pile of Leaves              

## Syncing with the Course Repository

Given all the issues we had in class, I think you should all reclone the course repository (unless everything is already working for you). Here's how to do it:

1. First delete the current 2024_RNA-Seq-Analysis RStudio projedct on your computer. You can do this by navigating to the project in your file explorer and deleting it. 

2. Open RStudio and select `File -> New Project -> Version Control -> Git`.

**Note**: If you don't have git installed on your device, you might have to first download and install it from [github](https://git-scm.com/downloads). Close Rstudio before installing git and re-open it afterwards.
	
3. In the "Repository URL" field, paste the following URL for the course material repository: `https://github.com/gurinina/2024_RNA-Seq-Analysis.git`

4. Choose a local folder on your computer to store the project files.

5. Click "Create Project" to clone the repository.

After the repository is cloned, you can sync your copy of 2024_RNA-Seq-Analysis with the GitHub course repository 2024_RNA-Seq-Analysis by clicking the "Pull" button in the "Git" tab in RStudio. This ensures that you have the latest version of the course materials.

## Installing Required R Packages

Once you have R, RStudio, and the course repository set up, you will need to install some R packages to work with the course material.

Copy and paste the following command into the RStudio console to install all the required packages:

```{r, results = "hide", message = FALSE, eval = FALSE}

remove.packages("GOenrichment")

# Install the "GOenrichment" package from GitHub
if (!requireNamespace("devtools", quietly = TRUE)) {
 install.packages("devtools")
}

devtools::install_github("gurinina/GOenrichment", force = TRUE)

# Install Bioconductor packages using BiocManager
if (!requireNamespace("BiocManager", quietly = TRUE)) {
 install.packages("BiocManager")
}

# Install required Bioconductor and CRAN packages
BiocManager::install(c(
 "bookdown", "clusterProfiler", "DESeq2", "dplyr", "enrichplot", "fgsea", "apeglm",
 "ggplot2", "ggrepel", "gplots", "knitr", "org.Hs.eg.db", "pheatmap", 
 "purrr", "RColorBrewer", "rmarkdown", "rsconnect", "tidyverse", "ggraph", "igraph", "visNetwork", "GO.db", "GOSemSim"
), force = TRUE)

```

## Verifying Your Installation

To verify that everything is installed correctly, load the packages by copying and pasting this into your RStudio console:

```{r, eval = FALSE}

library(knitr)
library(pheatmap)
library(bookdown)
library(DESeq2)
library(RColorBrewer)
library(tidyverse)
library(dplyr)
library(gplots)
library(ggrepel)
library(ggplot2)
library(GOenrichment)
library(GO.db)
library(fgsea)
library(enrichplot)
library(clusterProfiler)
library(org.Hs.eg.db)
library(purrr)
library(rmarkdown)
library(GOSemSim)
library(visNetwork)
library(igraph)
library(ggraph)

```

If the libraries load without any error messages, everything is set up correctly.

### Running R code; compiling the book

1. **Open the bookdown project:**

- Navigate to the `Files` pane in RStudio and open the `index.Rmd` file.
   
2. **Compile the book:**

- Copy and paste the following code in the R console to compile the book:

```{r, eval = FALSE}

# load the bookdown package
library(bookdown)

# removes all files from the _book  directory
clean_book(clean = TRUE)

# renders the book in GitBook format
bookdown::render_book("index.Rmd", "bookdown::gitbook")

# opens the GitBook in your browser
rstudioapi::viewer("_book/index.html")

# To see the book in your Viewer pane, open 'index.Rmd' in your main directory and click the 'Knit' button.
```

## Resyncing with the course repository

If you have changed any files in the course repository and want to get the latest version from GitHub, you can do so by following these steps:

1. **Open the Terminal in RStudio:**

- Navigate to the `Tools` menu and select `Terminal` > `New Terminal`.

2. **Type in the Commands:**

- In the terminal, run:
   
```bash

git fetch origin
git reset --hard origin/main
     
```

- These commands will update your 2024_RNA-Seq-Analysis copy to match the remote 2024_RNA-Seq-Analysis course `main` branch, discarding your changes.

- If you have any questions or need assistance with any part of the above, please let me know, and I will help you troubleshoot. I would like to hear from each of you to confirm that you have successfully knitted the bookdown file.

Please email me either way regardless of what computer you are using.

Thank you, Guri
