# GO
```{r, message=FALSE,warning=FALSE, echo=FALSE}

knitr::opts_chunk$set(
message = FALSE,
warning = FALSE,
comment = NA,
echo = TRUE
)

source("source/solutions.R")
library(ggplot2)
library(details)
library(GOenrichment)
library(tidyverse)

```

CLASS SOURCE

```{r}

print("classSOURCE")
```

```{r solutions-go-network, echo=FALSE}
solutions(expression(
  res_tableKD <- readRDS("data/res_tableKD.RDS"),
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble(),

  foldchanges <- res_tableKD_tb$log2FoldChange,
  names(foldchanges) <- res_tableKD_tb$gene,
  foldchanges <- sort(foldchanges, decreasing = TRUE),

  matx <- cbind(foldchanges, foldchanges),
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58),
  kresp <- runGORESP(fdrThresh = 0.2, scoreMat = scoreMat,
    bp_input = hGOBP.gmt, go_input = NULL, minSetSize = 20,
    maxSetSize = 300),

  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat),

  runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")
))


```


<div class="fold s o">
### Click to Show/Hide Solution
```{r , echo=FALSE}
solutions(expression(
  res_tableKD <- readRDS("data/res_tableKD.RDS"),
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble(),

  foldchanges <- res_tableKD_tb$log2FoldChange,
  names(foldchanges) <- res_tableKD_tb$gene,
  foldchanges <- sort(foldchanges, decreasing = TRUE),

  matx <- cbind(foldchanges, foldchanges),
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58),
  kresp <- runGORESP(fdrThresh = 0.2, scoreMat = scoreMat,
    bp_input = hGOBP.gmt, go_input = NULL, minSetSize = 20,
    maxSetSize = 300),

  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat),

  runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")
))
```
</div>




<div class="fold s o">
### Click to Show/Hide Solution
```{r, echo = FALSE}


```{r, class.source='fold-hide', echo=FALSE}
  res_tableKD <- readRDS("data/res_tableKD.RDS")
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble()
  
  foldchanges <- res_tableKD_tb$log2FoldChange
  names(foldchanges) <- res_tableKD_tb$gene
  foldchanges <- sort(foldchanges, decreasing = TRUE)
  
  matx <- cbind(foldchanges, foldchanges)
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58)
  
  kresp <- runGORESP(
    fdrThresh = 0.2, 
    scoreMat = scoreMat,
    bp_input = hGOBP.gmt, 
    go_input = NULL, 
    minSetSize = 20, 
    maxSetSize = 300
  )
  
  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat)
  
  runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")
```


NEETWORK ONLY FOLD SO
```{r}

print("Network OMLY")
```

<div class="fold s o">
### Network OMLY SO
```{r, echo = FALSE}

  runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")



```
</div>

<details markdown='1'><summary>Toggle</summary>
```{r,fig.show='hide'}


  res_tableKD <- readRDS("data/res_tableKD.RDS")
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble()
  
  foldchanges <- res_tableKD_tb$log2FoldChange
  names(foldchanges) <- res_tableKD_tb$gene
  foldchanges <- sort(foldchanges, decreasing = TRUE)
  
  matx <- cbind(foldchanges, foldchanges)
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58)
  
  kresp <- runGORESP(
    fdrThresh = 0.2, 
    scoreMat = scoreMat,
    bp_input = hGOBP.gmt, 
    go_input = NULL, 
    minSetSize = 20, 
    maxSetSize = 300
  )
  
  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat)
  
  v = runNetwork(kvis$nodes, kvis$edges, 
             main = "MOV10 Knockdown GO Enrichment Network")
  print(v)

```
</details>


EXPRESSION

```{r}

print("solutioEXPRESS")
```



```{r solutions-go-network, fig.show='hide'}



solutions(expression(
  res_tableKD <- readRDS("data/res_tableKD.RDS"),
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble(),

  foldchanges <- res_tableKD_tb$log2FoldChange,
  names(foldchanges) <- res_tableKD_tb$gene,
  foldchanges <- sort(foldchanges, decreasing = TRUE),

  matx <- cbind(foldchanges, foldchanges),
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58),
  kresp <- runGORESP(fdrThresh = 0.2, scoreMat = scoreMat,
    bp_input = hGOBP.gmt, go_input = NULL, minSetSize = 20,
    maxSetSize = 300),

  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat),

  v = runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")
  
))
print(v)
```

`

FOLD SO
```{r}



print("fold s o")
```

<div class="fold s">
### Click to SFOLD SO Solution
```{r , fig.show='hide'}


  res_tableKD <- readRDS("data/res_tableKD.RDS")
  res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble()

  foldchanges <- res_tableKD_tb$log2FoldChange
  names(foldchanges) <- res_tableKD_tb$gene
  foldchanges <- sort(foldchanges, decreasing = TRUE)

  matx <- cbind(foldchanges, foldchanges)
  scoreMat <- compSCORE(matx, coln = 1, sig = 0.58)
  
  kresp <- runGORESP(fdrThresh = 0.2, scoreMat = scoreMat,
    bp_input = hGOBP.gmt, go_input = NULL, minSetSize = 20,
    maxSetSize = 300)

  kvis <- visSetup(kresp$enrichInfo, kresp$edgeMat)
 v= runNetwork(kvis$nodes, kvis$edges, main = "MOV10 Knockdown GO Enrichment Network")
 print(v)


```
</div>


<div class="fold s">
### Click to SFOLD SO Solution
```{r ,fig.show='hide'}


res_tableKD <- readRDS("data/res_tableKD.RDS")

res_tableKD_tb <- res_tableKD %>% 
    data.frame() %>% rownames_to_column(var = "gene") %>% 
    dplyr::filter(!is.na(log2FoldChange)) %>% as_tibble()
    
  
kdfoldchanges <- res_tableKD_tb$log2FoldChange

names(kdfoldchanges) <- res_tableKD_tb$gene

kdfoldchanges <- sort(kdfoldchanges, decreasing = TRUE)

matkd <- cbind(kdfoldchanges,kdfoldchanges)

scoreMatkd = compSCORE(matkd, coln = 1, sig =  0.58)

kresp = runGORESP(fdrThresh = 0.2,bp_input = hGOBP.gmt,go_input = NULL,minSetSize = 20,maxSetSize = 300,scoreMat = scoreMatkd)


kvis = visSetup(kresp$enrichInfo,kresp$edgeMat)

v=runNetwork(kvis$nodes,kvis$edges)

print(v)
```
</div>
